Query Optimization for Dynamic Imputation

### 2017-03-01 paper/algorithm.tex
paper: system R -> canonical system, in most mentions

The space of query plans is similar to that considered by System R~\cite{blasgen1981system}, with the addition of imputation operators appearing before and after traditional operators.
\Cref{fig:query-schematic} shows a plan schematic for a query involving three tables, absent any imputation operators.

@@

As a reference implementation, \ProjectName{} uses a general-purpose imputation strategy
based on chained-equation decision trees.  Chained-equation
imputation methods~\cite{vanbuuren2011mice} (sometimes called \textit{iterative
regression}~\cite{gelman2006data}) impute multiple missing attributes by iteratively fitting
a predictive model of one missing attribute, conditional on all complete attributes
and all other missing attributes. Chained-equation decision trees algorithms are found to be
effective~\cite{akande2015empirical} and are widely used in epidemiological
domains~\cite{burgette2010multiple}. For our experiments, we use Weka's {\tt REPTree}
implementation (\cite{witten2016data}).

@@

These imputation algorithms are used as part of a larger \textit{multiple imputation}
process, in which multiple distinct copies of the complete data are generated and estimators
are computed by averaging over the multiple datasets to form a single completed dataset
(\cite{gelman2006data}). In \ProjectName{}, we consider only single imputation, though in
further work, multiple imputation could be another avenue of assessing
imputation confidence.

@@

\subsection{Complexity}\label{sec:complexity}
Our optimization algorithm builds off the approach taken by System R~\cite{blasgen1981system}.
If imputation operators are ignored, we search the same plan space.
Therefore, our algorithm has a complexity of at least $O(2^J)$, where $J$ is the number of joins.

>>>>

The space of query plans is similar to that considered in a canonical
cost-based optimizer~\cite{blasgen1981system}, with the addition of imputation operators appearing before and after traditional operators.
\Cref{fig:query-schematic} shows a plan schematic for a query involving three tables, absent any imputation operators.

@@

As a reference implementation, \ProjectName{} uses a general-purpose imputation strategy
based on chained-equation decision trees.  Chained equation
imputation methods~\cite{vanbuuren2011mice} (sometimes called \textit{iterative
regression}~\cite{gelman2006data}) impute multiple missing attributes by iteratively fitting
a predictive model of one missing attribute, conditional on all complete attributes
and all other missing attributes. Chained-equation decision trees algorithms are found to be
effective~\cite{akande2015empirical} and are widely used in epidemiological
domains~\cite{burgette2010multiple}. For our experiments, we use Weka's {\tt REPTree}
implementation (\cite{witten2016data}).

@@

These imputation algorithms are used as part of a larger \textit{multiple imputation}
process, in which multiple distinct copies of the complete data are generated and estimators
are computed by averaging over the multiple datasets to form a single completed dataset
(\cite{gelman2006data}). In \ProjectName{}, we consider only single imputation, though in
further work, multiple imputation could be considered as another avenue of assessing
imputation confidence.

@@

\subsection{Complexity}\label{sec:complexity}
Our optimization algorithm builds off the approach taken by 
a canonical cost-base optimizer~\cite{blasgen1981system}.
If imputation operators are ignored, we search the same plan space.
Therefore, our algorithm has a complexity of at least $O(2^J)$, where $J$ is the number of joins.
